Almost compatible functions and infinite length games

### 2015-12-10 articles/almost-compatible-functions-and-topological-games/article.tex
begin cleaning omega_2 cohen reals argument, including deflation of math mode

  But \(CH\) is not required to have \(\alcompS{\aleph_2}\) fail.

  % Alan's results

            \begin{lemma}[\cite{MR597342}]
            Assume that $G\subset \operatorname{Fn}(\omega_2,2)$ is a generic filter,
             and let $\mu\in \omega_2$. Then the final model $V[G]$ can be
             regarded as forcing  with $\operatorname{Fn}(\omega_2\setminus \mu,
             2)$ over the model $V[G_\mu]$.
            In addition, for each $\operatorname{Fn}(\omega_2,2)$-name $\dot A$
            of a subset of $\omega$ (treat as a subset of $\omega\times
            \operatorname{Fn}(\omega_2,2) $),
            there is a canonical name $\dot A(G_\mu)$ where,
            $$\dot A(G_\mu) = \{ (n,p\restriction [\mu,\omega_2))  :
             (n,p)\in \dot A\ \ \mbox{and} \ \ p\restriction \mu\in G_\mu\}$$
            and we get that the valuation of $\dot A(G_\mu)$ by the tail
            of the generic, $G_{\omega_2\setminus \mu}$, is the same as
            the valuation of $\dot A$ by the full generic.
            \end{lemma}

            \begin{theorem}
              There exists a model of \(ZFC\) for which \(\mf c=\aleph_2\)
              and \(\neg\alcompS{\aleph_2}\).
            \end{theorem}

            \begin{proof}
            The forcing poset is $\operatorname{Fn}(\omega_2, 2)$.
            Let $\{ \dot f_{A} : A\in [\omega_2]^\omega\}$ be a family of
            names such that $\dot f_{A}$ is a one-to-one function from $A$ into
            $\omega$. It suffices to only consider sets $A$ from the ground
            model.
            \bigskip

            Put all the lemma  stuff in an elementary submodel $M$ of the universe
            (technically of $H(\kappa)$,  or of $V_\kappa$,
             for some large $\kappa$). Standard methods says that we can assume that
             $|M|=\omega_1 =\mathfrak c$ and that $M^\omega\subset M$ (which means
             that every countable subset of $M$ is a member of $M$).
            \bigskip

            Let $\lambda = M\cap \omega_2$ (same as the supremum of $M\cap
            \omega_2$). Consider the name  $\dot f_{[\lambda,\lambda+\omega)}$.
            What is such a name?  We can assume that it is a set of pairs
             of the form $( (\lambda+k,m), p)$ where $p\in \mathop{Fn}(\omega_2,
             2)$ and, of course, $k,m\in \omega$. This is (almost) equivalent to
             saying
             that $p$ forces that $\dot f_{[\lambda,\lambda+\omega)}(\lambda+k) =
             m$. We don't take all such $p$, in fact for each $k,m$ it is enough
            to take a
             countable set of such $p$ to get an equivalent name
             (Kunen calls it a nice name if we take, for each $k,m$ an
            antichain that is maximal among such conditions).
            Given any such name $\dot f$, let $\operatorname{supp}(\dot f)$
            denote the union of the domains of conditions $p$ appearing in the
            name.

             Also let $Y$ equal $\operatorname{supp}(\dot
             f_{[\lambda,\lambda+\omega)})\setminus \lambda$.
             Let $\delta$ denote the order type of
             $Y$ and let the 2-parameter notation
             $\varphi_{\mu,\lambda}$ be the order-preserving function from
            $\mu\cup Y$ onto  the ordinal $\mu+\delta$.  This lifts canonically to
            an order-preserving bijection $\varphi_{\mu,\lambda}:
            \operatorname{Fn}(\mu\cup Y,2) \mapsto
            \operatorname{Fn}(\mu+\delta,2)$.
            Similarly, we make sense of the name
             $\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})$, call it $F_M$.
            Here simply, for each tuple $(~(k,m)~, p)\in \dot
            f_{[\lambda,\lambda+\omega)}$,
             we have that $(~(k,m)~,\varphi_{\mu,\lambda}(p))$ is in $F_M$.
            Again, let $\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})$
            be interpreted in the above sense as giving $F_M$ (which is an element
            of $M$).  Note that we do not regard $\delta$ as fixed here, but
            rather simply depending on the $\operatorname{supp}(\dot
            f_{[\lambda,\lambda+\omega)})$ described above. Other values
            replacing $\lambda>\mu$ will result in their own set $Y$
            and canonical map $\varphi_{\mu,\lambda}$; but one thing we do have to
            assume (or arrange) for other values $\alpha$ replacing $\lambda$
            is that $\operatorname{supp}(\dot f_{[\alpha,\alpha+\omega)})$
            intersected with
            $\alpha$ is contained in $\mu$.

            Now the object $F_M$ is an
            element of $M$, and $M$ believes this statement is true:
            $$
            (\forall \beta\in\omega_2)~ (\exists \beta<\lambda\in \omega_2)~~~
            \operatorname{supp}(\dot f_{[\lambda,\lambda+\omega)})
            \cap \lambda\subset \mu \ \ \mbox{and}\ \
            F_M = \varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})$$

            But now, this  means that,
            not only is there  an $\alpha\in M$,
            $ F_M = \varphi_{\mu,\alpha}(\dot f_{[\alpha,\alpha+\omega)})$
            but also that there is an increasing sequence
             $\{\alpha_\xi : \xi  \in \omega_1\}\subset \lambda$ of
             such $\alpha$'s satisfying that, for each $\xi$
            we have that $\operatorname{supp}(\dot
            f_{[\alpha_\xi,\alpha_\xi+\omega)})$ is contained in $\alpha_{\xi+1}$.

            Choose such a sequence.
             This means that if we let $A = \bigcup_{n>0} [\alpha_n,\alpha_n+\omega)$ we
             have the name $\dot f_A$ in $M$. This then means that all
             the $( (\beta,m) , p)$ appearing in $\dot f_A$ have the property
             that $\mathop{dom}(p)$ is contained in $M$.
            There is, within $M$, a name $\dot g$ satisfying that
             $\dot f_A(\alpha_n+k) = \dot f_{[\alpha_n,\alpha_n+\omega)}(\alpha_n+k)$
            for all $k >\dot g(n)$.

@@

            We now apply the above Lemma using $\mu = \mu_0$ and we are now
            working in the extension $V[G_\mu]$. We will abuse the notation
            and use $\dot f_{[\alpha_n,\alpha_n+\omega)}$ instead of
              $\dot f_{[\alpha_n,\alpha_n+\omega)}(G_{\mu})$ as defined in
              the Lemma.
            We work for a contradiction. Something special has now happened,
             namely,  the supports of the names $\{
            \dot f_{[\alpha_n, \alpha_n+\omega)} :  0< n<\omega\}$ are pairwise disjoint
            and also disjoint from the support of the name
             $\dot f_{[\lambda,\lambda+\omega)}$ (under the same convention about
             $G_\mu$.  And not only that, these names are pairwise isomorphic (in
             the way that they all map to $F_M$).

@@

            Since $A$ is disjoint from $[\lambda,\lambda+\omega)$,
            there must be an integer $\ell$
            together with a condition
             $q\in \mathop{Fn}(\omega_2\setminus \mu,2)$ satisfying that for all
             $n>\ell$ , $q$ forces that

            \centerline{
             ``if $k>\dot g(n)$ (since  $\alpha_n+k\in A$) then
             $\dot f_{[\alpha_n,\alpha_n+\omega)}(\alpha_n+k) \neq
            \dot f_{[\lambda,\lambda+\omega)}(\lambda +k)$''.}

            Choose $n$ large enough so that $\mathop{dom}(q) \cap [\alpha_n,
            \mu_{n+1})$ is
            empty.
             Choose $q_1<q\restriction \lambda$ (in $M$) so that
            $$ \varphi_{\mu,\alpha_n}(q_1\restriction \operatorname{supp}(
             \dot f_{[\alpha_n,\alpha_n+\omega)}) =
            \varphi_{\mu,\lambda}(q\restriction \operatorname{supp}(
             \dot f_{[\lambda,\lambda+\omega)}) $$
            and then (again in $M$) choose $q_2 < q_1$
            so that it both forces a value
             $L$ on $\ell+\dot g(n)$
            and subsequently forces a value $m$ on
             $\dot f_{[\alpha_n, \alpha_n+\omega)}(\alpha_n+L+1)$.
            But now, again calculate
            $$ q_3 = \varphi_{\mu,\lambda}^{-1} \circ \varphi_{\mu,\alpha_n}
            (q_2\restriction \operatorname{supp}(\dot
            f_{[\alpha_n,\alpha_n+\omega)}))$$
            and, by the isomorphisms, we have that $q_3$ forces that
             $\dot f_{[\lambda,\lambda+\omega)}(\lambda+L+1) = m$.

            Technically (or with more care)  all of this is taking place in the
            poset $\operatorname{Fn}(\omega_2\setminus \mu,2)$ and this means
            that $q_3$ and $q$ are all compatible with each other.

            Follow
            the bouncing ball: it suffices to consider
             $q(\beta)=e$ and to assume that $q_3(\beta)$ is defined.
            Since $q_3(\beta)$ is defined, we have that
            there is a  $\beta'\in\mathop{dom}(q_2)$ such that $
             \varphi_{\mu,\lambda}(\beta) = \varphi_{\mu,\alpha_n}(\beta')$,
            and that $q_3(\beta) = q_2(\beta')$.
            But, by definition of $q_1$, $\beta'\in\mathop{dom}(q_1)$
            and even  that $q_1(\beta') = q(\beta)$.
             Then, since $q_2<q_1$, we have that $q_2(\beta')=q_1(\beta') =
             q(\beta)$. This completes the circle that $q_3(\beta) = q(\beta)$.
            \bigskip

            Finally, our contradiction is that $q_3\cup q_2\cup q$
            forces that
             $k=L+1$ violates the quoted statement above.
            \end{proof}

>>>>

  But \(CH\) is not required to have \(\alcompS{\aleph_2}\) fail.
  We take the following lemma from Kunen.

  \begin{lemma}[\cite{MR597342}]
    Assume that $G\subset \operatorname{Fn}(\omega_2,2)$ is a generic filter,
    and let $\mu\in \omega_2$. Then the final model $V[G]$ can be
    regarded as forcing  with $\operatorname{Fn}(\omega_2\setminus \mu,
    2)$ over the model $V[G_\mu]$.
    In addition, for each $\operatorname{Fn}(\omega_2,2)$-name $\dot A$
    for $A\subseteq\omega$ (abusing notation and letting \(n\) represent
    its own $\operatorname{Fn}(\omega_2,2)$-name and
    $\operatorname{Fn}(\omega_2\setminus \mu, 2)$-name),
    there is a canonical name $\dot A(G_\mu)$ where,
    $$\dot A(G_\mu) = \{ (n,p\restriction [\mu,\omega_2))  :
    (n,p)\in \dot A\ \ \mbox{and} \ \ p\restriction \mu\in G_\mu\}$$
    and the valuation of $\dot A(G_\mu)$ by the tail
    of the generic, $G_{\omega_2\setminus \mu}$, is the same as
    the valuation of $\dot A$ by the full generic.
  \end{lemma}

 \begin{lemma}[\cite{MR597342}]
    Assume that \(G\subset \operatorname{Fn}(\omega_2,2)\) is a generic filter,
    and let \(\mu\in \omega_2\). Then the final model \(V[G]\) can be
    regarded as forcing  with \(\operatorname{Fn}(\omega_2\setminus \mu,
    2)\) over the model \(V[G_\mu]\).
    In addition, for each \(\operatorname{Fn}(\omega_2,2)\)-name \(\dot A\)
    for \(A\subseteq\omega\) (abusing notation and letting \(n\) represent
    its own \(\operatorname{Fn}(\omega_2,2)\)-name and
    \(\operatorname{Fn}(\omega_2\setminus \mu, 2)\)-name),
    there is a canonical name \(\dot A(G_\mu)\) where,
    \[\dot A(G_\mu) = \{ (n,p\restriction [\mu,\omega_2))  :
    (n,p)\in \dot A\ \ \mbox{and} \ \ p\restriction \mu\in G_\mu\}\]
    and the valuation of \(\dot A(G_\mu)\) by the tail
    of the generic, \(G_{\omega_2\setminus \mu}\), is the same as
    the valuation of \(\dot A\) by the full generic.
  \end{lemma}

  \begin{theorem}
    It follows from the consistency of \(ZFC\) that
    there exists a model of \(ZFC\) for which \(\mf c=\aleph_2\)
    and \(\neg\alcompS{\aleph_2}\).
  \end{theorem}

  \begin{proof}
    The forcing poset is \(\operatorname{Fn}(\omega_2, 2)\).
    Let \(\{ \dot f_{A} : A\in [\omega_2]^\omega\}\) be a family of
    names such that \(f_{A}\) is a one-to-one function from \(A\) into
    \(\omega\) in the ground model. Let \(M\) be an elementary submodel
    of the universe such that
    \(|M|=\omega_1 =\mathfrak c\) and \(M^\omega\subset M\).

    Let \(\lambda = M\cap \omega_2=\sup(M\cap\omega_2)\).
    Consider the name
    \[
      \dot f_{[\lambda,\lambda+\omega)}
        =
      \{
        ( (\lambda+k,m), p)
      :
        p\in \mathop{Fn}(\omega_2,2);k,m\in \omega
      \}
    \]
    This is (almost) equivalent to
             saying
             that \(p\) forces that \(\dot f_{[\lambda,\lambda+\omega)}(\lambda+k) =
             m\). We don't take all such \(p\), in fact for each \(k,m\) it is enough
            to take a
             countable set of such \(p\) to get an equivalent name
             (Kunen calls it a nice name if we take, for each \(k,m\) an
            antichain that is maximal among such conditions).
            Given any such name \(\dot f\), let \(\operatorname{supp}(\dot f)\)
            denote the union of the domains of conditions \(p\) appearing in the
            name.

             Also let \(Y\) equal \(\operatorname{supp}(\dot
             f_{[\lambda,\lambda+\omega)})\setminus \lambda\).
             Let \(\delta\) denote the order type of
             \(Y\) and let the 2-parameter notation
             \(\varphi_{\mu,\lambda}\) be the order-preserving function from
            \(\mu\cup Y\) onto  the ordinal \(\mu+\delta\).  This lifts canonically to
            an order-preserving bijection \(\varphi_{\mu,\lambda}:
            \operatorname{Fn}(\mu\cup Y,2) \mapsto
            \operatorname{Fn}(\mu+\delta,2)\).
            Similarly, we make sense of the name
             \(\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})\), call it \(F_M\).
            Here simply, for each tuple \((~(k,m)~, p)\in \dot
            f_{[\lambda,\lambda+\omega)}\),
             we have that \((~(k,m)~,\varphi_{\mu,\lambda}(p))\) is in \(F_M\).
            Again, let \(\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})\)
            be interpreted in the above sense as giving \(F_M\) (which is an element
            of \(M\)).  Note that we do not regard \(\delta\) as fixed here, but
            rather simply depending on the \(\operatorname{supp}(\dot
            f_{[\lambda,\lambda+\omega)})\) described above. Other values
            replacing \(\lambda>\mu\) will result in their own set \(Y\)
            and canonical map \(\varphi_{\mu,\lambda}\); but one thing we do have to
            assume (or arrange) for other values \(\alpha\) replacing \(\lambda\)
            is that \(\operatorname{supp}(\dot f_{[\alpha,\alpha+\omega)})\)
            intersected with
            \(\alpha\) is contained in \(\mu\).

            Now the object \(F_M\) is an
            element of \(M\), and \(M\) believes this statement is true:
            \[
            (\forall \beta\in\omega_2)~ (\exists \beta<\lambda\in \omega_2)~~~
            \operatorname{supp}(\dot f_{[\lambda,\lambda+\omega)})
            \cap \lambda\subset \mu \ \ \mbox{and}\ \
            F_M = \varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})\]

            But now, this  means that,
            not only is there  an \(\alpha\in M\),
            \( F_M = \varphi_{\mu,\alpha}(\dot f_{[\alpha,\alpha+\omega)})\)
            but also that there is an increasing sequence
             \(\{\alpha_\xi : \xi  \in \omega_1\}\subset \lambda\) of
             such \(\alpha\)'s satisfying that, for each \(\xi\)
            we have that \(\operatorname{supp}(\dot
            f_{[\alpha_\xi,\alpha_\xi+\omega)})\) is contained in \(\alpha_{\xi+1}\).

            Choose such a sequence.
             This means that if we let \(A = \bigcup_{n>0} [\alpha_n,\alpha_n+\omega)\) we
             have the name \(\dot f_A\) in \(M\). This then means that all
             the \(( (\beta,m) , p)\) appearing in \(\dot f_A\) have the property
             that \(\mathop{dom}(p)\) is contained in \(M\).
            There is, within \(M\), a name \(\dot g\) satisfying that
             \(\dot f_A(\alpha_n+k) = \dot f_{[\alpha_n,\alpha_n+\omega)}(\alpha_n+k)\)
            for all \(k >\dot g(n)\).

@@

            We now apply the above Lemma using \(\mu = \mu_0\) and we are now
            working in the extension \(V[G_\mu]\). We will abuse the notation
            and use \(\dot f_{[\alpha_n,\alpha_n+\omega)}\) instead of
              \(\dot f_{[\alpha_n,\alpha_n+\omega)}(G_{\mu})\) as defined in
              the Lemma.
            We work for a contradiction. Something special has now happened,
             namely,  the supports of the names \(\{
            \dot f_{[\alpha_n, \alpha_n+\omega)} :  0< n<\omega\}\) are pairwise disjoint
            and also disjoint from the support of the name
             \(\dot f_{[\lambda,\lambda+\omega)}\) (under the same convention about
             \(G_\mu\).  And not only that, these names are pairwise isomorphic (in
             the way that they all map to \(F_M\)).

@@

            Since \(A\) is disjoint from \([\lambda,\lambda+\omega)\),
            there must be an integer \(\ell\)
            together with a condition
             \(q\in \mathop{Fn}(\omega_2\setminus \mu,2)\) satisfying that for all
             \(n>\ell\) , \(q\) forces that

            \centerline{
             ``if \(k>\dot g(n)\) (since  \(\alpha_n+k\in A\)) then
             \(\dot f_{[\alpha_n,\alpha_n+\omega)}(\alpha_n+k) \neq
            \dot f_{[\lambda,\lambda+\omega)}(\lambda +k)\)''.}

            Choose \(n\) large enough so that \(\mathop{dom}(q) \cap [\alpha_n,
            \mu_{n+1})\) is
            empty.
             Choose \(q_1<q\restriction \lambda\) (in \(M\)) so that
            \[ \varphi_{\mu,\alpha_n}(q_1\restriction \operatorname{supp}(
             \dot f_{[\alpha_n,\alpha_n+\omega)}) =
            \varphi_{\mu,\lambda}(q\restriction \operatorname{supp}(
             \dot f_{[\lambda,\lambda+\omega)}) \]
            and then (again in \(M\)) choose \(q_2 < q_1\)
            so that it both forces a value
             \(L\) on \(\ell+\dot g(n)\)
            and subsequently forces a value \(m\) on
             \(\dot f_{[\alpha_n, \alpha_n+\omega)}(\alpha_n+L+1)\).
            But now, again calculate
            \[ q_3 = \varphi_{\mu,\lambda}^{-1} \circ \varphi_{\mu,\alpha_n}
            (q_2\restriction \operatorname{supp}(\dot
            f_{[\alpha_n,\alpha_n+\omega)}))\]
            and, by the isomorphisms, we have that \(q_3\) forces that
             \(\dot f_{[\lambda,\lambda+\omega)}(\lambda+L+1) = m\).

            Technically (or with more care)  all of this is taking place in the
            poset \(\operatorname{Fn}(\omega_2\setminus \mu,2)\) and this means
            that \(q_3\) and \(q\) are all compatible with each other.

            Follow
            the bouncing ball: it suffices to consider
             \(q(\beta)=e\) and to assume that \(q_3(\beta)\) is defined.
            Since \(q_3(\beta)\) is defined, we have that
            there is a  \(\beta'\in\mathop{dom}(q_2)\) such that \(
             \varphi_{\mu,\lambda}(\beta) = \varphi_{\mu,\alpha_n}(\beta')\),
            and that \(q_3(\beta) = q_2(\beta')\).
            But, by definition of \(q_1\), \(\beta'\in\mathop{dom}(q_1)\)
            and even  that \(q_1(\beta') = q(\beta)\).
             Then, since \(q_2<q_1\), we have that \(q_2(\beta')=q_1(\beta') =
             q(\beta)\). This completes the circle that \(q_3(\beta) = q(\beta)\).
            \bigskip

            Finally, our contradiction is that \(q_3\cup q_2\cup q\)
            forces that
             \(k=L+1\) violates the quoted statement above.
            \end{proof}
